---
title: "constraint function"
format:
  html:
    theme: cosmo
    code-fold: false
editor: visual
---

```{r, message = FALSE}
library(lubridate)
library(dplyr)
library(ggplot2)
```

## Simulation data

```{r}
set.seed(12345)

# --- Parameters ---
n_rows <- 300
survey_months <- c("2022-12-01", "2023-04-01", "2023-08-01", "2023-12-01")
survey_months <- as.Date(survey_months)

# Function to sample random date within a month
random_date_in_month <- function(start_date, n) {
  start <- as.Date(format(start_date, "%Y-%m-01"))
  end <- seq(start, by = "1 month", length.out = 2)[2] - 1
  as.Date(runif(n, as.numeric(start), as.numeric(end)), origin = "1970-01-01")
}


sigmoid <- function(x, beta0 = -3, beta1 = 0.5) {
  1 / (1 + exp(-(beta0 + beta1 * x)))
}

# --- Simulate data ---
simulated_data <- data.frame(
  time = unlist(lapply(survey_months, function(m) random_date_in_month(m, n_rows / length(survey_months)))),
  age = runif(n_rows, min = 0, max = 15) # random ages
)

# Probability of being seropositive follows a sigmoid in age
simulated_data$prob_sero <- sigmoid(simulated_data$age, beta0 = -3, beta1 = 0.5)

# Simulate binary outcome
simulated_data$seropositive <- rbinom(n_rows, size = 1, prob = simulated_data$prob_sero)

simulated_data <- simulated_data |> mutate(date = as.Date(time),
           col_time = rep(c(1:4),each = 75)) |> select(-c(time,prob_sero))
```

Assume the dataset has four columns: age (continuous variable), seropositive status (binary), date of collection, and collection time. This function models the age-seroprevalence profile as a monotonically increasing function of age for each collection time.

```{r}
head(simulated_data)
```

### Unconstrained age profile

```{r}
#| fig-width: 15
#| fig-height: 6
#| out-width: "100%"
#| fig-format: png

simulated_data %>%
ggplot(aes(x = age, y = seropositive)) +
  # geom_jitter(height = 0.05)+
  geom_point(aes(x = age, seropositive),
             shape = "|")+
  facet_wrap(~factor(col_time),
             ncol = 4)+
  geom_smooth(fill = "blue",alpha = 1/10,
              method = mgcv::gam,formula = y ~ s(x, bs = "bs"),
              method.args = list(method = "REML",link = "logit",
                                 family = "binomial"))+
  labs(y = "Seroprevalence (%)",x = "Age")+
  scale_x_continuous(breaks = seq(0,15,by = 3), minor_breaks = NULL)+
  scale_y_continuous(labels = scales::label_percent(), minor_breaks = NULL)+
  coord_cartesian(ylim = c(0, 1))+
  theme_bw()+
  theme(axis.title.x = element_text(size = 18),
        axis.text.x = element_text(size = 18),
        legend.position = "none",
        plot.tag = element_text(face = "bold", size = 18),
        axis.title.y = element_text(size = 18),
        axis.text.y = element_text(size = 18),
        strip.text.x = element_text(size = 18))
```

# Function

## Utils function

```{r}
age_profile <- function(data,age_values, ci = .95) {
  model <- gam(seropositive ~ s(age), binomial, data)
  link_inv <- family(model)$linkinv
  df <- nrow(data) - length(coef(model))
  p <- (1 - ci) / 2
  model |>
    predict(list(age = age_values), se.fit = TRUE) %>%
    c(list(age = age_values), .) |>
    as_tibble() |>
    mutate(lwr = link_inv(fit + qt(    p, df) * se.fit),
           upr = link_inv(fit + qt(1 - p, df) * se.fit),
           fit = link_inv(fit)) |>
    select(- se.fit)
}

shift_right <- function(n, x) {
  if (n < 1) return(x)
  c(rep(NA, n), head(x, -n))
}
```

## Main function

```{r,message=FALSE,warning=FALSE}
library(purrr)
library(tidyr)
library(magrittr)
library(mgcv)
library(scam)
```


```{r}
age_profile_constrained <- function(data,ci = .95,aging_correct = F){

  mean_collection_times <- data |>
    mutate(col_date2 = as.numeric(date)) |>
    group_by(col_time) |>
    summarise(mean_col_date = mean(col_date2)) |>
    with(setNames(mean_col_date, col_time))

  age_range <- range(data$age) %>% round()
  age_values <-  seq(age_range[1],age_range[2], le = 512)

  if (aging_correct == F){

  outcome <- data  |>
    group_by(col_time) |>
    group_modify(~ .x %>% 
                   age_profile(age_values = age_values,ci = ci) |>
                   mutate(across(c(fit, lwr, upr), ~ map(.x, ~ rbinom(100, 1, .x))))) |>
    ungroup() |>
    mutate(collection_time = mean_collection_times[as.character(col_time)]) |>
    unnest(c(fit, lwr, upr)) |>
    pivot_longer(c(fit, lwr, upr), names_to = "line", values_to = "seropositvty") |>
    group_by(age, line) |>
    group_modify(~ .x %>% 
                   scam(seropositvty ~ s(collection_time, bs = "mpi"), binomial, .) |>
                   predict(list(collection_time = mean_collection_times),type = "response") |>
                   as.vector()  %>% 
                   tibble(collection_time = mean_collection_times,
                          seroprevalence  = .)) |>
    ungroup() |>
    group_by(collection_time, line) |>
    group_modify(~ .x %>% 
                   mutate(across(seroprevalence, ~ gam(.x ~ s(age), betar) |>
                                   predict(type = "response") |> as.vector()))) |>
    ungroup() |>
    pivot_wider(names_from = line, values_from = seroprevalence) |>
    group_by(collection_time) |>
    group_split()

  } else {

  dpy <- 365

  cohorts <- cumsum(c(0, diff(mean_collection_times))) |>
      divide_by(dpy * mean(diff(age_values))) |>
      round() |>
      map(shift_right, age_values)

  age_time <- map2(mean_collection_times, cohorts,
                     ~ tibble(collection_time = .x, cohort = .y))

  age_time_inv <- age_time |>
      map(~ cbind(.x, age = age_values)) |>
      bind_rows() |>
      na.exclude()

  outcome <- data |>
    # Step 1:
    group_by(col_time) |>
    group_modify(~ .x  %>% 
                   age_profile(age_values = age_values, ci = ci) |>
                   mutate(across(c(fit, lwr, upr), ~ map(.x, ~ rbinom(100, 1, .x))))) |>
    group_split() |>
    map2(age_time, bind_cols) |>
    bind_rows() |>
    unnest(c(fit, lwr, upr)) |>
    pivot_longer(c(fit, lwr, upr), names_to = "line", values_to = "seropositvty") |>
    # Step 2a:
    filter(cohort < max(age) - diff(range(mean_collection_times)) / dpy) |>
    group_by(cohort, line) |>
    group_modify(~ .x %>% 
                   scam(seropositvty ~ s(collection_time, bs = "mpi"), binomial, .) |>
                   predict(list(collection_time = mean_collection_times),type = "response") |> as.vector() %>% 
                   tibble(collection_time = mean_collection_times,
                          seroprevalence  = .)) |>
    ungroup() |>
    # Step 2b:
    left_join(age_time_inv, c("cohort", "collection_time")) |>
    group_by(collection_time, line) |>
    group_modify(~ .x %>% 
                   right_join(tibble(age = age_values), "age") |>          ### added
                   arrange(age) |>                                         ### added
                   mutate(across(seroprevalence,
                                 ~ gam(.x ~ s(age), betar) |>
                                   predict(list(age = age_values),type = "response") %>% as.vector()))) %>%   ### modified
    ungroup() |>
    pivot_wider(names_from = line, values_from = seroprevalence) |>
    group_by(collection_time) |>
    group_split()

  }

  return(outcome)
}
```

```{r,warning=FALSE}
aaa <- age_profile_constrained(simulated_data,ci = .95,aging_correct = T)

aaa2 <- age_profile_constrained(simulated_data,ci = .95,aging_correct = F)
```

## Age profile constrained without aging correct

```{r}
#| fig-width: 10
#| fig-height: 8
#| out-width: "100%"
#| fig-format: png

aaa2 %>% 
  bind_rows() %>%
  ggplot(aes(x = age, y = fit,
                 color = factor(collection_time,labels = c("1","2","3","4")), fill = factor(collection_time,labels = c("1","2","3","4")))) +
  geom_ribbon(aes(ymin = lwr, ymax = upr), alpha = 0.2, color = NA) +
  geom_line() +
  labs(x = "age (years)", y = "sero-prevalence", color = "Group", fill = "Group") +
  theme_minimal(base_size = 14)+
  theme(legend.position = "bottom")
```

## Age profile constrained with aging correct

```{r}
#| fig-width: 10
#| fig-height: 8
#| out-width: "100%"
#| fig-format: png

aaa %>% 
  bind_rows() %>%
  ggplot(aes(x = age, y = fit,
                 color = factor(collection_time,labels = c("1","2","3","4")), fill = factor(collection_time,labels = c("1","2","3","4")))) +
  geom_ribbon(aes(ymin = lwr, ymax = upr), alpha = 0.2, color = NA) +
  geom_line() +
  labs(x = "age (years)", y = "sero-prevalence", color = "Group", fill = "Group") +
  theme_minimal(base_size = 14)+
  theme(legend.position = "bottom")
```

## apply for HFMD serodata

```{r, warning=FALSE,message=FALSE}
library(stringr)
```


```{r}
hfmd_sero <- readRDS("D:/OUCRU/hfmd/data/hfmd_sero.rds")

hfmd <- hfmd_sero %>%
  as_tibble() |>
  mutate(collection = id |>
           str_remove(".*-") |>
           as.numeric() |>
           divide_by(1e4) |>
           round(),
         col_date2 = as.numeric(col_date),
         across(pos, ~ .x > 0))
```

## Age profile constrained without aging correct

```{r}
hfmd_seroo <- hfmd %>% rename(seropositive = pos,
                col_time = collection,
                date = col_date) %>%
  age_profile_constrained(ci = .95,aging_correct = F)

hfmd_seroo %>% bind_rows() %>%
  ggplot(aes(x = age, y = fit,
             color = factor(collection_time), fill = factor(collection_time))) +
  geom_ribbon(aes(ymin = lwr, ymax = upr), alpha = 0.2, color = NA) +
  geom_line() +
  labs(x = "age (years)", y = "sero-prevalence", color = "Group", fill = "Group") +
  theme_minimal(base_size = 14)+
  theme(legend.position = "hide")
```

## Age profile constrained with aging correct

```{r}

hfmd_seroo2 <- hfmd %>% rename(seropositive = pos,
                              col_time = collection,
                              date = col_date) %>%
  age_profile_constrained(ci = .95,aging_correct = T)

hfmd_seroo2 %>% bind_rows() %>%
  ggplot(aes(x = age, y = fit,
             color = factor(collection_time), fill = factor(collection_time))) +
  geom_ribbon(aes(ymin = lwr, ymax = upr), alpha = 0.2, color = NA) +
  geom_line() +
  labs(x = "age (years)", y = "sero-prevalence", color = "Group", fill = "Group") +
  theme_minimal(base_size = 14)+
  theme(legend.position = "hide")
```


